Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2019-07-31 paper/manuscript.tex
prerequisites fixes

%%%%%%%%%%%%
% Step 0   %
%%%%%%%%%%%%
\section{What can be expected from alchemical simulations? -- Step 0}
\label{sec:step0}
\todo[inline, color={green!20}]{@everyone please review this section}
Step 0 is essentially about initially counting the cost of your
project: Can you even hope to tackle the problem you are thinking
about attempting given available resources and, if successful, will it
be worth it?
\subsection*{How accurate are current alchemical free energy calculations?}
Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative~\cite{lots-of-citations}. 
\todo[inline,color={green!20}]{JDC: We should probably provide a short list of key benchmark references here, broken down by force field, and keep this up to date. GT: would a table, at the end of the document be an easier way to do this, only requiring some brief comments in the text?}
However, the domain of applicability is a significant concern~\cite{Sherborne2016, Cournia2017}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

It is worth noting that the accuracy of free energy calculations is highly variable across different protein targets, and likely across different ligand chemotypes as well. For instance, FEP+ with OPLS3 achieves an RMSE of 0.62 kcal/mol for a set of 21 compounds binding to JNK1 kinase, yet only an RMSE of 1.05 kcal/mol for a set of 34 compounds binding to P38$\alpha$ kinase (cite Harder et al. 2016). Furthermore, perturbations for the same chemotype in different pockets of the BACE enzyme gave varied errors (cite keranen et al jctc 2017, 13 1439)  
\todo[inline,color={green!20}]{JDC: It's important to be clear on what error is being reported: $\Delta G$ after shifting by a constant to minimize the RMSE, unshifted $\Delta G$, $\Delta \Delta G$ of computed edges, or $\Delta \Delta G$ of all edges. Note that computed edge $\Delta \Delta G$ RMSE significantly underestimates the error of all edge RMSE: \url{https://github.com/jchodera/jacs-dataset-analysis}}
We
thus recommend that retrospective study for a particular target and a particular chemical series be performed to establish the relevant accuracy of free energy calculations for each particular application case.

\subsection{Is the system of study suitable for alchemical free energy calculations?}
Before even planning free energy calculations to study binding to a
particular target, it is important to assess what is known about the
system and its timescales and its suitability for free energy
calculations, as well as the \emph{purpose} of the calculations and
the amount of available computer resources. In some cases, predicting
accurate binding free energies for a particular target might be a
\emph{more} challenging effort than simply measuring them! This is
often the case when dealing with database screening problems, where
compounds might be easily and quickly available commercially for
testing and free energy calculations could consume a much larger
amount of resources. Free energy calculations thus typically only
appeal when (slow or costly) synthesis would be required to do
experiments or experiments are otherwise cost-prohibitive.

@@

\subsection{Is the expected accuracy of the computation good enough?}
Additionally, the requisite level of accuracy is important -- if the
goal is to guide lead optimization when many compounds will be
synthesized, free energy calculations can be appealing even with
accuracies in the 1-2 kcal/mol range [cite Brown/Mobley/Shirts,
  Klimovich/Mobley Perspective], but if the number of compounds to be
synthesized is very small, this accuracy may not be enough to provide
much value.

@@

\todo[inline, color={green!20}]{@DM: A plot of this might be useful. I believe you have used/generated one of these before.}

@@

prohibitive, or at least should be carefully considered. And, are

@@

\subsection*{Is an explorative study what I want?}

@@

\todo[inline, color={blue!20}]{@Gary, here may be a good spot to add some lessons learned from FEP in drug discovery.  }
%
%
%
%%%%%%%%%%%%
% Step 1   %
%%%%%%%%%%%%
%
\section{Simulation prerequisites -- Step 1}
\label{sec:step1}
\todo[inline,color={green!20}]{@everyone: please review}
The alchemical free energy protocols, which are discussed in Section~\ref{sec:step2}, which can be used are tied in with what type of free energy you might want to compute, i.e. a free energy of binding or a free energy of hydration. Depending on the type of simulations different considerations have to be made for ligands, solvent, and host molecules (in the case of the estimation of free energies of binding).
\subsection*{Free energies of binding}
In principle in the limit of sufficient conformational sampling, the free energy changes estimated from an alchemical free energy calculation should be independent of the system's initial coordinates. However, in practice because simulations of thermodynamic states are of finite duration (typically currently 1-100 ns per state) this is only true for certain classes of alchemical free energy calculations such as relative or absolute free energies of hydration of small and relatively rigid organic molecules. Host-guest or protein-ligand complexes typically exhibit slowly relaxing degrees of freedom that exceed significantly the duration of an alchemical free energy calculation. It is therefore generally important to carefully select input coordinates to obtain satisfactory results. 
You might want to ask yourself the following questions before diving into your simulation setup. 
\begin{itemize}
    \item Do I have one or multiple good host structures? (e.g. a good resolution X-ray crystal of the protein target)
    \item Should I include burried waters, or other small molecules that can be found in an X-ray structure.
    \item Are my ligands part of a congeneric series? (simple R group substitutions around the same scaffold)
    \item Do I have information on one or all of the ligands binding sites (e.g. a X-ray structure)
\end{itemize}

>>>>

%%%%%%%%%%%%
% Step 0   %
%%%%%%%%%%%%
\section{What can be expected from alchemical simulations? -- Step 0}
\label{sec:step0}
\todo[inline, color={green!20}]{@everyone please review this section}
Step 0 is about initially counting the cost of your
project: Can you even hope to tackle the problem you are thinking
about attempting given available resources and, if successful, will it
be worth it?
\subsection*{How accurate are current alchemical free energy calculations?}
Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative~\cite{lots-of-citations}. 
\todo[inline,color={green!20}]{JDC: We should probably provide a short list of key benchmark references here, broken down by force field, and keep this up to date. GT: would a table, at the end of the document be an easier way to do this, only requiring some brief comments in the text? ASJSM: I like the table idea.}
However, the domain of applicability is a significant concern~\cite{Sherborne2016, Cournia2017}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

It is worth noting that the accuracy of free energy calculations is highly variable across different protein targets, and likely across different ligand chemotypes as well. For instance, FEP+ with OPLS3 achieves an RMSE of 0.62 kcal/mol for a set of 21 compounds binding to JNK1 kinase, yet only an RMSE of 1.05 kcal/mol for a set of 34 compounds binding to P38$\alpha$ kinase~\cite{harder2016opls3}. Furthermore, perturbations for the same chemotype in different pockets of the BACE enzyme gave varied errors~\cite{keranen2017acylguanidine}
\todo[inline,color={green!20}]{JDC: It's important to be clear on what error is being reported: $\Delta G$ after shifting by a constant to minimize the RMSE, unshifted $\Delta G$, $\Delta \Delta G$ of computed edges, or $\Delta \Delta G$ of all edges. Note that computed edge $\Delta \Delta G$ RMSE significantly underestimates the error of all edge RMSE: \url{https://github.com/jchodera/jacs-dataset-analysis} ASJSM: Can we incorporate your analysis in this section as an example?}
We thus recommend that retrospective study for a particular target and a particular chemical series be performed to establish the relevant accuracy of free energy calculations for each particular application case.
%
\subsection*{Is the system of study suitable for alchemical free energy calculations?}
Before even planning free energy calculations to study binding to a
particular target, it is important to assess what is known about the
system and its timescales and its suitability for free energy
calculations, as well as the \emph{purpose} of the calculations and
the amount of available computer resources. In some cases, predicting
accurate binding free energies for a particular target might be a
\emph{more} challenging effort than simply measuring them! This is
often the case when dealing with database screening problems, where
compounds might be easily and quickly available commercially for
testing and free energy calculations could consume a much larger
amount of resources. Free energy calculations thus typically only
appeal when (slow or costly) synthesis would be required to do
experiments or experiments are otherwise cost-prohibitive.

@@

\subsection*{Is the expected accuracy of the computation good enough?}
Additionally, the requisite level of accuracy is important -- if the
goal is to guide lead optimization when many compounds will be
synthesized, free energy calculations can be appealing even with
accuracies in the 1-2 kcal/mol range~\cite{brown2010freeenergy,mobley2012perspective}, but if the number of compounds to be
synthesized is very small, this accuracy may not be enough to provide
much value.

@@

\todo[inline, color={blue!20}]{@DM: A plot of this might be useful. I believe you have used/generated one of these before.}

@@

prohibitive, or at least should be carefully considered. 
Furthermore, are

@@

\subsection*{Is an exploratory study what I want?}

@@

%
%
%
%%%%%%%%%%%%
% Step 1   %
%%%%%%%%%%%%
%
\section{Simulation prerequisites -- Step 1}
\label{sec:step1}
\todo[inline,color={green!20}]{@everyone: please review}
The alchemical free energy protocols, which are discussed in Section~\ref{sec:step2}, which can be used are tied in with what type of free energy you might want to compute, i.e. a free energy of binding or a free energy of hydration. Depending on the type of simulations different considerations have to be made for ligands, solvent, and host molecules (in the case of the estimation of free energies of binding).
\subsection*{Free energies of binding}
In principle in the limit of sufficient conformational sampling, the free energy changes estimated from an alchemical free energy calculation should be independent of the system's initial coordinates. However, in practice because simulations of thermodynamic states are of finite duration (typically currently 1-100 ns per state) this is only true for certain classes of alchemical free energy calculations such as relative or absolute free energies of hydration of small and relatively rigid organic molecules. Host-guest or protein-ligand complexes typically exhibit slowly relaxing degrees of freedom that exceed significantly the duration of an alchemical free energy calculation. It is therefore generally important to carefully select input coordinates to obtain satisfactory results. 
You might want to ask yourself the following questions before diving into your simulation setup. 
\begin{itemize}
    \item Do I have one or multiple good host structures? (e.g. a good resolution X-ray crystal of the protein target)
    \item Should I include burried waters, or other small molecules that can be found in an X-ray structure.
    \item Are my ligands part of a congeneric series? (simple R group substitutions around the same scaffold)
    \item Do I have information on one or all of the ligands binding sites (e.g. a X-ray structure)
\end{itemize}
